---
title: 20INMCAL204- Laboratory Report
author:
  - name: Naveen Prince
    email: naveen.prince@saintgits.org
    affiliation: Student, INMCA 20-25
    correspondingauthor: true
    footnote: 1
  - name: Siju K.S
    email: siju.swamy@saintgits.org
    affiliation: Saintgits College of Engineering (Autonomous)
    footnote: 2
  - name: Jobin Jose
    email: jobin.jose@saintgits.org
    affiliation: Saintgits College of Engineering (Autonomous)
    footnote: 2
address:
  - code: Student, INMCA 20-25
    address: Department of Computer Applications, RB402
  - code: Saintgits College of Engineering (Autonomous)
    address: Department of Mathematics, AB304
footnote:
  - code: 1
    text: "Student, 20INMCAL204."
  - code: 2
    text: "Course Faculty, 20INMCAL204."
abstract: |
 Experiments listed in the Lab Manual are successfully executed in the R version 4.1.0. Details of the experiments with input \& ouput are summerized in the form of a report. Experiments are arranged in the form of sections. This report is prepared using the R-package `rticles` [@mrticles].
journal: "Mathematics Department for Evaluation"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: mybibfile.bib
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
---
\tableofcontents
\newpage
# Experiment 4: Statistical Summary and measure of normality of a dataset

## Aim

1. To create the statistical summary of a data

2. To study normality of the data

## Packages used and syntax of R methods

For statistical summary of a given dataset, the `rbase` package will be used. To calculate skewness and kurtosis of dataset, the `ACSWR` is used.

>*Note:* The functions `skewness` and `kurtosis` from the `e1071` package are more generic functions. Another resouse is `moments` package. 

## Algorithm
* Step 1: Load the dataset

* Step 2: Load necessary packages

* Step 3: Calculate statistical summaries

* Step 4: Calculate the `skewness` and `kurtosis` of the numerical data

* Step 5: Report the results

## R code

```{r}
#loading package
library(ACSWR)
#loading data
data(yb)
#view structure of data
str(yb)

```
```{r}
# creating statistical summary

summary(yb)
```
```{r}
range(yb$Preparation_1); range(yb$Preparation_2) # list out ranges of data

```

```{r}
#skewness and kurtosis of preparation_1
skewcoeff(yb$Preparation_1); kurtcoeff(yb$Preparation_1)
```
```{r}
#skewness and kurtosis of preparation_2
skewcoeff(yb$Preparation_2); kurtcoeff(yb$Preparation_2)
```

## Results & discussions

A distribution is normal then `mean=median=mode` and the skewness is 0 and kurtosis is 2. In this experiment statistical summaries of two variables are created. From the skewness and kurtosis measures, both the variables are positively skewed and `preparation_1` is lepto-kurtic and `preparation_2` is meso-kurtic. Based on the statistical summary and skewness and kurtosis measures, both the variables are different from a normal distribution.

# Experiment 5- Implementation of Bayes Theorem

## Aim

1. To calculate Bayes posterior probability using Bayes theorem

## Packages used and syntax of R methods

Bayes posterior probability can be directly calculated using mathematical method or using the package `LaplacesDemon`.

## Algorithm
* Step 1: Load the package, prior probabilities and conditionals

* Step 2: Calculate the Bayes posterior probability using the formula-$P(B_j|A)=\dfrac{P(A|B_j)P(B_j)}{\sum\limits_{j=1}^mP(A|B_j)P(B_j)}$

* Step 3: Calculate the same prior probability using `LaplaceDemon` package

* Step 4: Report the results

>*Case:* Classical Problem from Hoel, Port, and Stone (1971). Suppose there are three tables with two drawers each. The first table has a gold coin in each of the drawers, the second table has a gold coin in one drawer and a silver coin in the other drawer, while the third table has silver coins in both of the drawers. A table is selected at random and a drawer is opened which shows a gold coin. 

>*Observation:*The problem is to compute the probability of the other drawer also showing a gold coin. The Bayes formula can be easily implemented in an R program.

## R code

```{r}
#loading data
prob_GC <- c(1,1/2,0)
priorprob_GC <- c(1/3,1/3,1/3)
```

```{r}
#calculating postrior probability
post_GC <- prob_GC*priorprob_GC
post_GC/sum(post_GC)
```

```{r}
# do the same using LaplacesDemon` package
library(LaplacesDemon)
BayesTheorem(prob_GC, priorprob_GC)

```


## Results & discussions

The Bayes theorem is used to calculate posterior probability of the Mathematical model of the given case. Also the result is verified using the `LaplacesDemon` package.


\hrule

# References {.unnumbered}
