---
title: "Experiment-3 Descriptive Statistical Analysis"
author: "ss"
date: "22/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Aim

To administer baseline statistical analysis on a dataset and report descriptive analysis summary.


## Algorithm

* Step-1: Load the data and required R-packages for data analysis

* Step-2: Apply basic statistic functions 

* Step-3: Create appropriate visualizations

* Step-4: Report the findings based on descriptive analysis

## R-code

>Loading data

```{r}
df<-read.csv("https://raw.githubusercontent.com/sijuswamy/StatLab/main/Dataset_1.csv",header = TRUE)
df$Gender=as.factor(df$Gender)

head(df)
```

```{r}
str(df)
```
> Finding Column sums

```{r,warning=FALSE}
library(dplyr)
df1=select(df,-Student_Name,-Gender)
Sub_total=colSums(df1)
Sub_average=colMeans(df1)
```
```{r}
round(Sub_average,2)
```
```{r}
library(ggplot2)
crop=ggplot(data=df, mapping=aes(x=Gender, y=X20IMCAT201))+geom_boxplot()+labs(x ="Gender", y = "Maths")
crop
```

```{r}
ggplot(data = df, aes(x = X20IMCAT201,fill = df$Gender)) + geom_histogram(binwidth = 15, fill = "seagreen",color = "red")+ 
  theme(legend.position = "top")#+facet_grid(~Gender)

```
```{r}
plot(density(df$X20IMCAT201))
```


```{r}
median(df$X20IMCAT201)
```



```{r}
library(DescTools)
Mode(df$X20IMCAT201)
```
>User defined funtion


```{r}
calcmode <- function(a) {  
vector <- unique(a)  
vector[which.max(tabulate(match(a, vector)))]  
}  
```
```{r}
calcmode(df$X20IMCAT201)
```
```{r}
sd(df$X20IMCAT201)
```


## Results and discussions