---
title: 'Experiment 8: Normal Distribution'
author: "siju.swamy@saintgits.org"
date: "4/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Aim

1. To calculate probability mass density, probability distribution and quantiles using Normal distribution

## Packages used and syntax of R methods

Functions from `stat` package (which is loaded by default). 

The probability mass at a point $x$ can be evaluated using the syntax:

>`dnorm(x=x,mean=m,sd=s)`.

The probability distribution $P(X\leq x)$ is calculated using the `pnorm()` function. Syntax is: 

>`pnorm(x,mean=m,sd=s)`

The quantile for probability $p$ can be evaluated using the `qnorm()` function. Syntax is:

>`qnorm(prob,mean=m,sd=s)`

## Algorithm

* Step 1: Assign the inputs for required distribution

* Step 2: Calculate the required probabilities

* Step 3: Report the results


>*Case:* Generate and draw the cdf and pdf of a normal distribution with mean=10 and standard deviation=3. Use values of $x$ from 0 to 20 in intervals of 1.

## R code

```{r}
# create input parameters
t=seq(0,20,1); mu=10;sd=3

```
### Prbability distribution

```{r}
#calculating probability mass distribution and cumulative distribution
pmval=dnorm(t,mean = mu,sd=sd)
pmval
```

### Cumulative probability distribution

```{r}
#calculating cumulative density
cdval=pnorm(t,mean = mu,sd=sd)
cdval
```
### Plotting the `pmf` and `cdf`

```{r}
par(mfrow=c(1,2))
plot(t,pmval,xlab="t",ylab="P(X=t)", main="The Normal Distribution")
plot(t,cdval,xlab="t",ylab=expression(P(X<=t)),main="Cumulative Distribution Function")

```


## Results & discussions

The `pmf` and `cf` of the Normal distribution for given input parameters are evaluated and create respective plots.
